Polymethylation scores paper

### 2021-10-26 Papers/Polymethylation_scores/06_discussion.Rmd
edits to results and scripts for 8/25 FF meeting

Our analysis answers pertinent questions for the development of a clinically applicable methylation biomarker of prenatal maternal smoking. We show that DNA methylation scores calculated from children's saliva can accurately classify prenatal maternal smoke exposure in non-European ancestry children. Other bomarkers and risk scores are known to differ across ancestry groups, potentially contributing to and exacerbating health disparities. As new biomarkers are developed, understanding their consistency across ancestry groups is critical. We also demonstrated that accuracy is not substantially affected when using coefficients from newborn cordblood (no cell type control, age 9: 0.62, age 15: 0.61) vs older children peripheral blood (no cell type control age 9: 0.63, age 15: 0.60). However, using coefficients from models that incorporated cell-type control at the time of calculation did result in better accuracy, suggesting cell-type control should be instituted both when creating weights and when applying scores to new population. Including samples from both ages did not improve accuracy, thus repeated samplings may be unnecessary for clinical applications. Additionally, accuracy was more consistent when using polymethylation scores than using the single top-hit from Joubert et al, AHHR:CpG05575921 [@Joubert:2016ue]. Salivary AHHR:CpG05575921 methylation was less accurate in our sample at predicting prenatal maternal smoke exposure than when it has been used to classify own smoking in blood or saliva [@Philibert2020]. While using a single CpG site as a biomarker may be clinically pragamatic, single sites may be less robust to time-since-exposure and new environmental exposures. Alternatively, incorporating information from across several CpGs in a polymethylation score may be more robust to these .

>>>>

Our analysis answers pertinent questions for the development of a clinically applicable methylation biomarker of prenatal maternal smoking. We show that DNA methylation scores calculated from children's saliva can accurately classify prenatal maternal smoke exposure in non-European ancestry children. Other bomarkers and risk scores are known to differ across ancestry groups, potentially contributing to and exacerbating health disparities. As new biomarkers are developed, understanding their consistency across ancestry groups is critical. We also demonstrated that accuracy is not substantially affected when using coefficients from newborn cordblood (no cell type control, age 9: 0.62, age 15: 0.61) vs older children peripheral blood (no cell type control age 9: 0.63, age 15: 0.60). However, using coefficients from models that incorporated cell-type control at the time of calculation did result in better accuracy, suggesting cell-type control should be instituted both when creating weights and when applying scores to new population. Including samples from both ages did not improve accuracy, thus repeated samplings may be unnecessary for clinical applications. Additionally, accuracy was more consistent when using polymethylation scores than using the single top-hit from Joubert et al, AHRR:CpG05575921 [@Joubert:2016ue]. Salivary AHRR:CpG05575921 methylation was less accurate in our sample at predicting prenatal maternal smoke exposure than when it has been used to classify own smoking in blood or saliva [@Philibert2020]. While using a single CpG site as a biomarker may be clinically pragamatic, single sites may be less robust to time-since-exposure and new environmental exposures. Alternatively, incorporating information from across several CpGs in a polymethylation score may be more robust to these .

### 2021-10-26 Papers/Polymethylation_scores/index_paper.Rmd
edits to results and scripts for 8/25 FF meeting

@@

```{r child = here::here("Papers", "Polymethylation_scores", "05_results.Rmd")}
```

>>>>

```{r, echo=F, include=F, warning=F, message=F}
library(stringr)
results_run=str_detect(here::here(), 'nfs/turbo/bakulski1')
```

@@

```{r child = here::here("Papers", "Polymethylation_scores", "05_results.Rmd"), eval=results_run}
```
